---
title: "Homework 10: Decisions and loops"
author: "Katelyn Houghton"
format: gfm
editor: visual
---

# Decisions

Write a function that accepts the current time as a parameter and prints "Good morning", "Good afternoon", or "Good evening" depending on the time. It is fine for the time to be in numeric format (e.g., `2317` for 11:17 pm). Bonus points if the function accepts time objects (see the `lubridate` package).

```{r, message=FALSE}


```

A few questions to consider:

-   What is a logical name for this function? What is a logical name for the parameter it accepts?

    **Answer**:

-   The purpose of this function is to print a message to the console, so its primary purpose is a **side effect**. However, all functions must return something. What would be a logical value for this function to return?

    **Answer**:

-   Should the function have default behavior in case the user does not pass an argument?

    **Answer**:

-   What would you like to happen if this function is passed the wrong data type (e.g., a negative number)?

    **Answer**:

# Loops

-   Write a `for` loops that calculates the mean of each column of mtcars

```{r, message=FALSE}


```

-   Write a function (using a for loop) that calculates the mean of all numeric columns of *any* data frame. This function should be able to accept data frames with non-numeric columns.

```{r, message=FALSE}


```

## Why not loops

In R, we generally encourage people to use vectorized functions instead of `for` loops. According to [your textbook](https://r4ds.had.co.nz/iteration.html), what is better about vectorized functions?

**Answer**:
