---
title: "Homework 3"
author: "Katelyn"
format: gfm
editor: visual
---

# Data input/output

Use the tidyverse `read_csv()` function to read a .csv file of your own data. If you don't have any data of your own, make up an excel file with fake data.

```{r, message=FALSE}

library(tidyverse)

#Load my data file 
pups <- read.csv("hmk_03_data.csv")

```

# Investigating the properties of data frames

Use two different functions that both give some kind of summary or overview of the data frame. Which one do you think is more useful?

```{r, message=FALSE}

head(pups)

```

```{r, message=FALSE}

glimpse(pups)

```

I found the #head function more helpful because it arranged the summary in a table style that was easy to view. Also we can use "summary" or "str"

```{r, message=FALSE}

summary(pups)

```

```{r, message=FALSE}

str(pups)

```

# Manipulating data frames

Create a new column of your data frame by performing some kind of arithmetic operation on an existing column.

```{r, message=FALSE}

Fluff.percent <- pups$Fluff.factor *10

cbind(pups, Fluff.percent)

```

# Working on columns

Calculate the mean of a numeric column in your data frame.

```{r, message=FALSE}

mean(pups$Fetch.skills)

```

# Accessing elements of data frames

-   [Challenge 3 from SC Data Structures lesson](https://swcarpentry.github.io/r-novice-gapminder/04-data-structures-part1/index.html#challenge-3): Show what each of the following returns, and explain what each line of code is doing: \*\*Doing this instead for my pups data frame
    -   `pups[1] this will call the first column, and lay it out top to bottom`
    -   `pups[[1]] this will call the first column, but list the rows`
    -   `pups$Age this calls the data from the Age column`
    -   `pups[1, 1] this calls the data in the first row and column position`
    -   `pups[ , 1] this calls everything in the first column, but leaving a specific row blank`
    -   `pups[1, ] this calls the first row`

```{r, message=FALSE}

pups[1]
pups[[1]]
pups$Age
pups[1, 1]
pups[ , 1]
pups[1, ]

```

# Optional challenge

A data frame is an example of a list. I haven't discussed how to access elements of lists, but it is covered [here](https://swcarpentry.github.io/r-novice-gapminder/04-data-structures-part1/index.html#lists)).

Explain in what ways accessing elements of lists are like accessing columns of data frames, and given that, how it shows that data frames are a type of list.

# Subsetting our dataframe

https://swcarpentry.github.io/r-novice-gapminder/06-data-subsetting/index.html

We can subset using logical operations

```{r, message=FALSE}

pups[c(TRUE, FALSE, TRUE, TRUE) , ]

```

We can subset using logical test of dogs that are ten years old

```{r, message=FALSE}

pups[pups$Age == 10, ]

```

Select dogs that are seven years or older

```{r, message=FALSE}

pups[pups$Age >= 7, ]

```
